---
name: Hendro
surname: Sugandi
position: "PhD candidate in Finance at the Stockholm School of Economics"
www: "hendrosugandi.com"
email: "hendro.sugandi@phdstudent.hhs.se"
phone: +46 764272125
date: "`r format(Sys.time(), '%B %Y')`"
docname: CV
output: vitae::moderncv
header-includes:
 - \usepackage{savesym}
 - \savesymbol{fax}
 - \usepackage{marvosym} 
 - \restoresymbol{MARV}{fax}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

## headcolor: 414141
## headcolor: 990000
## 414141 
## "00008B"

require(xfun)

packages <- c('vitae'     # The CV's LaTeX Template
             ,'tibble'    # Data structure for tribbles
             ,'magrittr'  # The Pipe %>%
             ,'here'      # Relative File Paths
             ,'readr'     # Data Loading
             ,'glue'      # String Manipulation
            )

xfun::pkg_attach2(packages, message = FALSE)

library(tibble)

```

# Research interest

Empirical asset pricing, investments, and statistical learnings

<!-- market anomalies,  -->

# Education

```{r EDUCATION}

tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "PhD in Finance", "08/2016-(Expected) 2022", "Stockholm School of Economics (SSE)", "Stockholm, Sweden",
  "MSc in Financial Engineering", "08/2014-05/2016", "University of Illinois", "Urbana-Champaign, USA",
  "BSc in Industrial Engineering", "08/2006-08/2010", "Bandung Institute of Technology (ITB)", "Bandung, Indonesia"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)


```

# Job market paper

**Information uncertainty and investors’ expectations: evidence from earnings announcements**

**Abstract**: This paper investigates how information uncertainty amplifies investors’ expectation errors. I use earnings announcement periods to observe the change in investors’ expected firms’ payoffs. I show that investors are too optimistic (pessimistic) for stocks with low (high) expected earnings. The errors in expectations are large (small) in stocks with high (low) information uncertainty. Given a high level of information uncertainty, analysts’ estimates tend to be inaccurate and imprecise, and investors’ beliefs deviate from analysts’ forecasts.   


# Work in progress

- Target leverage and the cross-section of stock returns (joint with Michael Halling)
- Mutual funds and market anomalies
<!-- \vspace{12pt} -->

<!-- \pagebreak -->

# Seminars and conferences

<!-- - 10th National PhD Workshop in Finance (2021) -->
<!-- - Brown bag seminar at the Stockholm School of Economics (2021) -->
<!-- - PhD seminars at the Stockholm School of Economics (2019, 2020, 2021) -->

```{r SEMINARS}

tribble(
  ~ Year, ~ Presentation,
  "2021", "10th National PhD Workshop in Finance",
  "2021", "Brown bag seminar at the Stockholm School of Economics ",
  "2019, 2020, 2021", "PhD seminars at the Stockholm School of Economics "
) %>% 
  brief_entries(
      when = Year, 
      what = Presentation)

```

# Teaching experience at the Stockholm School of Economics (as a teaching assistant)

```{r Teaching}

# tribble(
#   ~ Year, ~ Teaching,
#   "2021", "Portfolio Choice and Asset Pricing (MSc in Finance)",
#   "2019, 2020, 2021", "Advanced Methods in Finance (MSc in Finance)",
#   "2019", "Risk Management (MSc in Finance)",
#   "2018", "Finance (Executive MBA, SSE)",
#   "2017", "Financial Management (Executive MBA, SSE)"
# ) %>% 
#   detailed_entries(
#       when = Year, 
#       what = Teaching)

tribble(
  ~ Year, ~ Teaching,
  "2021", "Portfolio Choice and Asset Pricing (MSc in Finance)",
  "2019, 2020, 2021", "Advanced Methods in Finance (MSc in Finance)",
  "2019", "Risk Management (MSc in Finance)",
  "2018", "Finance (Executive MBA)",
  "2017", "Financial Management (Executive MBA)"
) %>% 
  brief_entries(
      when = Year, 
      what = Teaching)
```


# Awards and scholarships

```{r Awards}

tribble(
  ~ Year, ~ Grant,
  "2021-2022", "Jan Wallander's and Tom Hedelius's Foundation, and Tore Browald's Foundation grant",
  "2018-2021", "The Swedish Bank Research Foundation (BFI) grant",
  "2016-2018", "Stockholm School of Economics Graduate Scholarship",
  "2014-2016", "The Indonesian Education Scholarship",
  ) %>% 
  brief_entries(
      when = Year, 
      what = Grant)


```


# Work experience

```{r}
tribble(
  ~ Year, ~ Work,
  "08/2018-10/2018", "Intern at research department at Sveriges Riksbank (Central Bank of Sweden)",
  "03/2011-01/2013", "Consultant at AJCapital Advisory",
  ) %>% 
  brief_entries(
      when = Year, 
      what = Work)

```


# Languages and computer skills

- Languages: Indonesian (native), English (fluent)
- Computer skills: R (advanced), Python, Matlab, Stata (intermediate/basic)   

# Other information

- Guest speaker for finance 101 sharing sessions: Indonesian Student Association in Stockholm/Sweden (2016, 2018, 2019), DWP at the Embassy of the Republic of Indonesia in Stockholm (2020), Indonesian Professionals in Sweden (2021), DWP at the Ministry of Agriculture, Indonesia (2021)
- Head of Industrial Engineering Student Association at ITB (2009-2010)
- Born: 1988
- Citizenship: Indonesian

\pagebreak

# References

\begin{tabular}{lr}
% Referee 1
\begin{minipage}[t]{3.5in}
Michael Halling (primary advisor)\\
Associate Professor\\
Stockholm School of Economics\\
\Letter\ \href{mailto:Michael.Halling@hhs.se}{Michael.Halling\textrm{@}hhs.se}\\
\Telefon\ +43 664 9114784\\
\end{minipage}
&
% Referee 2
\begin{minipage}[t]{3.5in}
Jungsuk Han\\
Associate Professor\\
Stockholm School of Economics\\
\Letter\ \href{mailto:jungsuk.han@hhs.se}{Jungsuk.Han\textrm{@}hhs.se}\\
\Telefon\ +46 8 736 91 58\\
\end{minipage}
\\
\\ % Additional newline for spacing.
% Referee 3
\begin{minipage}[t]{3.5in}
Riccardo Sabbatucci\\
Assistant Professor\\
Stockholm School of Economics\\
\Letter\ \href{mailto:Riccardo.Sabbatucci@hhs.se}{Riccardo.Sabbatucci\textrm{@}hhs.se}\\
\Telefon\ +46 8 736 92 93\\
\end{minipage}
\end{tabular}

```{r}
## https://tex.stackexchange.com/questions/34881/references-section-in-a-cv

```



